Distributed computation of linear algebra operations

### 2023-10-30 latex/presentation/main.tex
project reformat
determinant presentation

\documentclass{beamer}
\usetheme{metropolis}
\usepackage{verbatim}

@@

\title{Distributed computation of linear algebra operations}
\subtitle{Distributed systems and networks laboratory}
\author{
    Gabriele Aloisio [503264] \and
    Samuel Giacomo Raffa [matricola]
    }

@@

\section{Implementation in Python}
\begin{frame}{The Matrix and RayMatrix classes}
    Before implementing the parallelized algorithms for Ray, we wrote the code for the serial execution in the \alert{Matrix} class:
\end{frame}

\begin{frame}[fragile]{The Matrix class}
    \begin{verbatim}
        class Matrix:
            def __init__(self, elements):
                [...]
                self.elements = elements
            def inv(self): [...]
            def det(self): [...]
            def product(self, b): [...]
            def rank(self, order): [...]
            [...]

    \end{verbatim}

@@

\begin{frame}{The RayMatrix class}
    The \alert{RayMatrix} class is an extension of the Matrix class that inherits its attributes and methods. Operations that had to be parallelized were overrided and new auxiliary methods (methods that start with \textit{task\textunderscore}) decorated with \textbf{@ray.remote} were added:
\end{frame}

\begin{frame}[fragile]{The RayMatrix class}
    \begin{tiny}
        \begin{verbatim}
        class RayMatrix(Matrix):
            def __init__(self, elements):
                super().__init__(elements)

            @ray.remote
            def task_rank_det(submatrix, j):

            @ray.remote
            def task_get_square_submatrix(self, start_row, start_col, row, order):

            @ray.remote
            def task_det(self, elements, i):

            @ray.remote
            def task_inv_cof(self, a, i, j):

            @staticmethod
            @ray.remote
            def task_multiply(a, b, i, j, k):

            @staticmethod
            @ray.remote
            def task_sum(results):

    \end{verbatim}
    \end{tiny}
\end{frame}

@@

\begin{frame}[fragile]{Multiplication with Ray (1)}
    \begin{small}
        \begin{verbatim}
@staticmethod
def product(a, b):
    a_columns = a.size()["columns"]
    a_rows = a.size()["rows"]

    b_columns = b.size()["columns"]

    if a_rows != b_columns:
        raise ValueError(
            """
            Number of columns of the first matrix 
            must match the number of rows of the second matrix
            """
            )
    else:
    ...

    \end{verbatim}
    \end{small}
\end{frame}

\begin{frame}[fragile]{Multiplication with Ray (2)}
    \begin{tiny}
        \begin{verbatim}
a_elements = a.get()  # Get the elements of matrix a
b_elements = b.get()  # Get the elements of matrix b

result = [[0] * b_columns for _ in range(a_rows)]  # Create a result matrix filled with zeros
tasks = []  # Initialize an empty list to store the tasks

# Iterate over rows of matrix a
for i in range(a_rows):
    # Iterate over columns of matrix b
    for j in range(b_columns):
        elements_to_multiply = []  # Initialize an empty list to store the elements to be multiplied

        # Iterate over columns of matrix a and rows of matrix b
        for k in range(a_columns):
            # Add a task to multiply the elements at position (i, k) and (k, j)
            elements_to_multiply.append(
                RayMatrix.task_multiply.remote(a=a_elements, b=b_elements, i=i, j=j, k=k)
            )

        # Add the task to compute the sum of multiplied elements
        tasks.append(RayMatrix.task_sum.remote(results=ray.get(elements_to_multiply)))
    \end{verbatim}
    \end{tiny}
\end{frame}

\begin{frame}[fragile]{Multiplication with Ray (3)}
    \begin{verbatim}
# Get the results of all the tasks
results = ray.get(tasks)

# Fill the result matrix with the computed results
for i in range(a_rows):
    for j in range(b_columns):
        result[i][j] = results.pop(0)

return RayMatrix(result)

    \end{verbatim}
\end{frame}

\begin{frame}[fragile]{Multiplication with Ray(4)}
    \begin{verbatim}
        @staticmethod
        @ray.remote
        def task_multiply(a, b, i, j, k):
            return a[i][k] * b[k][j]
    \end{verbatim}
\end{frame}

@@

\begin{frame}[fragile]{Determinant with Ray (1)}
    \begin{tiny}
        \begin{verbatim}
def det(self):
if self.is_square():
    size = self.size()["rows"]
    a = self.get()

@@

            \end{verbatim}
    \end{tiny}
\end{frame}

\begin{frame}[fragile]{Determinant with Ray (2)}

    \begin{small}
        \begin{verbatim}
@ray.remote
def task_det(self, elements, i):
    size = len(elements)
    submatrix_det_sum = 0
    mats = self.get_square_submatrices(i)

    for j in range(1, size):
        submatrix_det = mats[j].det2()
        submatrix_det_sum += 
            elements[i][j] * ((-1) ** (i + j + 2)) 
                * submatrix_det

    return submatrix_det_sum
    \end{verbatim}
    \end{small}

@@

\begin{frame}[fragile]{Inverse with Ray}
    \begin{small}
        \begin{verbatim}
            codice inversa
        \end{verbatim}
    \end{small}
\end{frame}

@@

\begin{frame}[fragile]{Rank with Ray}
    \begin{small}
        \begin{verbatim}
            codice rango
        \end{verbatim}
    \end{small}
\end{frame}

>>>>

\documentclass{beamer}
\usetheme{metropolis}
\usepackage{float}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{transparent}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
% \definecolor{backcolour}{rgb}{0.95,0.95,0.95}
\definecolor{backcolour}{rgb}{0.98,0.98,0.98}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\fontsize{7}{5}\selectfont
    aboveskip=15pt, % Adjust the space above the listing
    belowskip=15pt, % Adjust the space below the listing
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=1
}

\lstset{style=mystyle}

@@

\title{Distributed computation of linear algebra operations}
\subtitle{Distributed systems and networks laboratory}
\author{Gabriele Aloisio [503264]}

@@

\section{Implementation in Python}
\begin{frame}{The Matrix classe}
    Before implementing the parallelized algorithms for Ray, we wrote the code for the serial execution in the \alert{Matrix} class:
\end{frame}

\begin{frame}[fragile]{The Matrix class}
    \begin{lstlisting}[language=Python]
class Matrix:
    def __init__(self, data):
        if isinstance(data, list):
            self.data = [[round(val, 8) for val in row]
                                        for row in data]
        else:
            raise ValueError("Input must be a list")
    [...]
    def dot(A, B): [...]
    def det(self): [...]
    def rank(self): [...]
    def inv(self): [...]
\end{lstlisting}

@@

\begin{frame}{The Matrix class}
    We modify the Matrix class in a later time to parallelize its operations
\end{frame}

@@

\begin{frame}[fragile]{Multiplication with Ray (Serial implementation)}
    \begin{lstlisting}[language=Python]
@staticmethod
def dot(A, B):
    if A.is_matrix() and B.is_matrix():
        a_rows, a_cols = A.shape()
        b_rows, b_cols = B.shape()

        if a_cols == b_rows:
            result = [[0] * b_cols for _ in range(A.shape()[0])]
            for i in range(A.shape()[0]):
                for j in range(b_cols):
                    for k in range(a_cols):
                        result[i][j] += A.data[i][k] * B.data[k][j]
            return Matrix(result)
        else:
            raise ValueError(
                "Matrix dimensions do not match for dot product")
    else:
        raise ValueError("Dot product requires a Matrix object")
    \end{lstlisting}
\end{frame}

\begin{frame}[fragile]{Multiplication with Ray (dot\_calc)}
    To parallelize this function, we simply delegate the calculation to an auxiliary function \textit{dot\_calc}:

    \begin{lstlisting}[language=Python]
@staticmethod
@ray.remote
def task_multiply(a, b, i, j, k):
    return a[i][k] * b[k][j]
    \end{lstlisting}

\begin{frame}[fragile]{Multiplication with Ray (Parallel)}
    \begin{lstlisting}[language=Python]
@staticmethod
def dot(A, B):
    if A.is_matrix() and B.is_matrix():
        a_rows, a_cols = A.shape()
        b_rows, b_cols = B.shape()

        if a_cols == b_rows:
            result = [[0] * b_cols for _ in range(a_rows)]
            futures = []

            for i in range(a_rows):
                for j in range(b_cols):
                    for k in range(a_cols):
                        futures.append(((i, j), t.dot_calc.remote(A, B, i, j, k)))

            for future in futures:
                i, j = future[0]
                result[i][j] += ray.get(future[1])

            return Matrix(result)
        else:
            raise ValueError(
                "Matrix dimensions do not match for dot product")
    else:
        raise ValueError("Dot product requires a Matrix object")
\end{lstlisting}
\end{frame}

@@

\begin{frame}[fragile]{Determinant with Ray (Serial)}
    \begin{lstlisting}[language=Python]
def det(self):
if self.is_square():
    size = self.size()["rows"]
    a = self.get()

@@

            \end{lstlisting}
\end{frame}

\begin{frame}[fragile]{Determinant with Ray (minor)}
    The \textit{minor} function
    \begin{lstlisting}[language=Python]
    def minor(self, i, j):
    '''
    Extract a minor matrix by removing the ith row and jth column
    '''

    data = self.get()
    minor_data = [row[:j] + row[j + 1:]
                    for row_idx, row in enumerate(data) if row_idx != i]

    return Matrix(minor_data)
\end{lstlisting}
    Then becomes the parallelized function \textit{dist\_minor}:
\end{frame}

\begin{frame}[fragile]{Determinant with Ray (dist\_minor)}
    \begin{lstlisting}[language=Python]
@ray.remote
def dist_minor(A, i, j):
    '''
    Extract a minor matrix by removing the ith row and jth column
    '''

    data = A.get()
    minor_data = [row[:j] + row[j + 1:]
                    for row_idx, row in enumerate(data) if row_idx != i]

    return Matrix(minor_data)
        \end{lstlisting}
    So finally the distributed determinant function will be:
\end{frame}

\begin{frame}[fragile]{Determinant with Ray (Parallel)}

    \begin{lstlisting}[language=Python]
    def det(self):
        if self.is_square():
            data = self.get()
            _, cols = self.shape()

            if cols == 1:
                return data[0][0]

            elif cols == 2:
                return (data[0][0] * data[1][1]) - (data[0][1] * data[1][0])

            else:
                det_value = 0

                minors_futures = [Matrix.dist_minor.remote(self, 0, j) for j in range(cols)]
                minors = ray.get(minors_futures)

                for minor, j in zip(minors, range(cols)):
                    # minor = self.minor(0, j)
                    det_value += ((-1) ** j) * data[0][j] * minor.det()

                return det_value
    \end{lstlisting}
\end{frame}

@@

\begin{frame}[fragile]{Inverse with Ray}
    \begin{lstlisting}[language=Python]
            codice inversa
        \end{lstlisting}
\end{frame}

@@

\begin{frame}[fragile]{Rank with Ray}
    \begin{lstlisting}[language=Python]
            codice rango
        \end{lstlisting}
\end{frame}
